---
title: "Project to Predict the Manner of Excercise"
author: "Ahn"
date: "Saturday, June 18, 2016"
output:
  html_document:
    theme: united
---

This is an assignment for the Machine Learning course of Coursera. 

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement. 
One of the information is about barbell lifts. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.The data is from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

The goal of your project is to predict the manner("classe" variable) in which they did the exercise. 

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. More information about data is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Modeling Process

## 1. Getting data

```{r}
trainingURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainingURL), na.strings=c("NA","#DIV/0!",""))
test <- read.csv("D:/Bigdata/Coursera/08. Machine Learning/pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

## 2. Exploring & Cleaning Data

Before modeling the data, we should explore data. 

```{r}
summary(training)
```

As an exploring data, some columns contained only NA values, so these were removed from the dataset. Using the nearZeroVar() function, I can find variables that were considered uninformative (zero or near zero variance predictors). Then, I make a new file without NZV variables. 
```{r}
library(caret)
training_NZV <- nearZeroVar(training, saveMetrics=TRUE)
training_names <- names(training)
training_NZV <- cbind(training_NZV, training_names)
training_NZV_2 <- subset(training_NZV, subset=(nzv==TRUE))
NZV_names <- training_NZV_2$training_names
NZV_names2 <- names(training) %in% NZV_names
training_2 <- training[!NZV_names2]
```

Some variables have too many NAs(90% and more). For Variables that have more than a 90% threshold of NA¡¯s I¡¯m going to leave them out:
```{r}
training_NA <- sapply(training_2, function(x) sum(is.na(x)/nrow(training_2)))
training_NA <- as.data.frame(training_NA)
training_names_2 <- names(training_2)
training_NA <- cbind(training_names_2, training_NA)
training_NA_2 <- subset(training_NA, subset=(training_NA>0.9))
NA_names <- training_NA_2$training_names_2
NA_names2 <- names(training_2) %in% NA_names
training_3 <- training_2[!NA_names2]
```


## 2. Modeling

After exploring data, I delete the first 7 values which are not related with the sensor measurements for modeling.
```{r}
new_train <- training_3[,-(1:7)]
```

I try to make 4 models with rpart, random forest, gbm and lda and will choose one based on accuracy and sample error. The object variable is multi-norminal, so I can't use 'GLM'method.

```{r}
library(rpart)
library(randomForest)
fit_rpart <- train(classe~., data=training, method="rpart")
fit_rf <- train(classe~., data=training, method="rf", prox=TRUE)

fit_rf <- train(classe~., data=training, method="rf", 
               trControl = trainControl(method="cv"), number=3)
fit_gbm <- train(classe~., data=training, method="gbm")
fit_lda <- train(classe~., data=training, method="lda",
                trControl=trainControl(method="cv"), number=3)
```





You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
